---
title: "STAT429 HW4"
author: "Taiga Hasegawa(taigah2)"
date: "2018/11/5"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Question1 

```{r}
rec = scan("cmort.dat")
data=rec
y=mean(data)
data_new=data-y
rec.yw=ar.yw(data_new,order=2)
n=length(rec)
```
The mean value of x is $\mu =E(X_t)=$ `y`

```{r}
rec.yw$ar
phi1=rec.yw$ar[1]
phi2=rec.yw$ar[2]
```
These are the point estimates of $\phi_1$ and $\phi_2$.

```{r}
rec.yw$var.pred
```

This is the point estimate of $\sigma^2$.

```{r}
sd=sqrt(diag(rec.yw$asy.var.coef))
stat1=(phi1-0.4)/sd[1]
stat2=(phi2-0.4)/sd[2]
2*pnorm(abs(stat1),lower.tail = FALSE)
2*pnorm(abs(stat2),lower.tail = FALSE)
```
Both p-value is over 30%. This means that we cannot reject the null hypothesis. The data is compatible with the hypothesis.

##Question 4

```{r}
P=4
Q=4
crit1=matrix(0,P+1,Q+1)
crit2=matrix(0,P+1,Q+1)
for (j in 0:P)
{
for (k in 0:Q)
{
dataML=arima(data,order=c(j,0,k),method="ML")
#AICC
crit1[j+1,k+1]=n*log(dataML$sigma)+2*(j+k+1)*n/(n-j-k-2)
#BIC
crit2[j+1,k+1]=n*log(dataML$sigma)+(j+k+1)*log(n)

}
}

```

```{r}
crit1
crit2
```
The best model is ARIMA(3,0,1) because the BIC is the smallest when the model is ARIMA(3,0,1)

```{r}
rec.ML<-arima(data,order=c(3,0,1),method="ML")
tsdiag(rec.ML,gof.lag=20)
```
If you see this, you can know that ACF of Residuals is between the blue line and p-values for Ljung-Box statistics are all above the blue line. This means that this model has almost 0 ACF of residuals. In addition we can be in favor of the fact that autocorrelations of data is zero. 

