---
title: "STAT429 HW06"
author: "Taiga Hasegawa(taigah2)"
date: "2018/12/12"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Question3 
```{r}
library(astsa)
data(sunspotz)
per= spec.pgram(sunspotz,taper=0, log="no")
1/(which(per$spec==sort(per$spec,decreasing = TRUE)[1])/per$n.used)
1/(which(per$spec==sort(per$spec,decreasing = TRUE)[2])/per$n.used)
1/(which(per$spec==sort(per$spec,decreasing = TRUE)[3])/per$n.used)
1/(which(per$spec==sort(per$spec,decreasing = TRUE)[4])/per$n.used)
```
There are 4 predominant periods, 22, 20, 23, 160 but the third one is not predominant period in the periodogram. Then I omit the third one and conclude that 22, 20, 160 is the predominant cycle. Because each year has 2 observations, we can also say that the predominat periods are 10, 11, and 80. 

```{r}
U = qchisq(.025,2) 
L = qchisq(.975,2) 
2*per$spec[22]/L 
2*per$spec[22]/U 
2*per$spec[24]/L 
2*per$spec[24]/U 
2*per$spec[3]/L 
2*per$spec[3]/U 
```

The confidence interval for period 22 is (8804.265,1282807). That for period 20 is (8290.672,1207975). That for period 160 is (3428.087,499482.4).

##Question4
```{r}
k=kernel("daniell",c(7,7))
l=kernel("daniell",c(14,14))
ave_1= spec.pgram(sunspotz, k, taper=0, log="no")
ave_2= spec.pgram(sunspotz, l, taper=0, log="no")
1/(which.max(ave_1$spec)/ave_1$n.used)
1/(which.max(ave_2$spec)/ave_2$n.used)
```
When spans=c(7,7), the predominant period is 22. When spans=c(14,14), the predominant period is 23. 
```{r}
 df_1 = ave_1$df   
 U_1 = qchisq(.025,df_1) 
 L_1 = qchisq(.975,df_1) 
df_1*ave_1$spec[22]/L_1
df_1*ave_1$spec[22]/U_1
```
Confidence interval when spans=c(7,7) is (5076.224,11945.4).
```{r}
 df_2 = ave_2$df   
 U_2 = qchisq(.025,df_2) 
 L_2 = qchisq(.975,df_2) 
df_2*ave_2$spec[21]/L_2
df_2*ave_2$spec[21]/U_2
```
Confidence interval when spans=c(14,14) is (3706.613,6832.477).
Spectral estimate is geting smoother and smoother as the frequency band is larger. 

##Question4
```{r}
acf(sunspotz,lag.max=50)
pacf(sunspotz)
```
The ACF and PACF shows that this model is AR(2).
```{r}
rec.ML<-arima(sunspotz,order=c(2,0,0),method="ML")
rec.ML
```
This is the estimation of coefficients of AR(2).

```{r}
armaspec=function(T=5000,arcoeff,macoeff,vare){

## Input
## T = resolution in frequency (length of time series)
## arcoeff = vector of coefficients of the AR component of the time series
## macoeff = vector of coefficients of the MA component
## vare = variance of the white noise error

## Output 
## specout = spectrum of the ARMA process defined from (0, pi]

## Note: This is the parameterization of the ARMA(p,q) process 
## that I will adopt in this program 
## X(t) = arcoeff[1]X(t-1) + ... + arcoeff[p]X(t-p) + 
## W(t) + macoeff[1]W(t-1) + ... + macoeff[q]W(t-q)
## I will assume that the process is already causal and 
## invertible, i.e., the user has done these checks before
## attempting to derive the spectrum.

freq = 2*pi*seq(0, T-1)/T;

arcoeff = c(1, -arcoeff);
macoeff = c(1, macoeff);

theta=c(1:T); phi=c(1:T);

for (k in 1:(T)){

theta.temp=0;
for (m in 1: (length(macoeff)))
{
theta.temp = theta.temp + macoeff[m]*complex(mod=1, arg=-1*m*freq[k])}
theta[k] = (abs(theta.temp))^2

phi.temp=0;
for (m in 1: (length(arcoeff))){
phi.temp = phi.temp + arcoeff[m]*complex(mod=1, arg=-1*m*freq[k])}
phi[k] = (abs(phi.temp))^2

}

spec = vare*(theta)/(phi)
specout = spec[2:(T/2 + 1)]
return(specout)
}
```

```{r}
truespec=armaspec(arcoeff=c(1.69,-0.788), macoeff=c(0), vare=1)
1/(which.max(truespec)/length(truespec))
```
The main periodicity is 11. 
```{r}
plot(1:length(truespec),truespec,col=2)
```
Autoregressive spectral estimator is much smoother than the conventional nonparametric estimator but the distribution is alike. 



